### 2017-03-28 thesis/intro/intro.tex
tmp commit

Obesity has been one of the major global problems for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
In fact, the risk of comorbidities increases with the increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.
The number of overweight and obese people, in both adults and children, have risen in every country of the world and the trend will likely to continue in the future.
Estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.

@@

Obesity is defined as an abnormal or excessive fat accumulation that may impair health of that individual \citep{Garrow1988}.
One common and widely used approach to categorise individuals is by measuring the \gls{bmi} of the individual.
\gls{bmi} is a measurement based on the weight-to-height ratio of an individual that is often used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight or obese categories.
The unit of \gls{bmi} is defined as the weight in kilograms per square of the height in meters (kg/m$^2$).

\gls{who} (2014) have used \gls{bmi} measurements to define overweight and obesity in adults as individuals with \gls{bmi} $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$, respectively.
For children under the age of 5, overweight and obesity are defined as weight-for-height greater than 2 standard deviation and 3 standard deviation above \gls{who} Child Growth Standards median, respectively.
For children aged between 5 to 19 years, overweight and obesity are defined as \gls{bmi}-for-age greater than 1 standard deviation and 2 standard deviation above \gls{who} Growth Reference Standards median, respectively.

However, the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the metabolic state of an individual remains controversial.
In fact, many studies have suggested the use of other measurements such as waist-to-height and waist-to-hip ratios to classify the patients into different categories, as these measurements were better predictors of the diseases related to obesity than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the \gls{bmi} information.

@@

From the latest global status report on noncommunicable diseases by \citet{WHO2014}:
\begin{quote}
	\textit{
		In 2014, 39\% of adults aged 18 years and older (38\% of men and 40\% of women) were overweight.
		The worldwide prevalence of obesity nearly doubled between 1980 and 2014.
		In 2014, 11\% of men and 15\% of women worldwide were obese.
		Thus, more than half a billion adults worldwide are classed as obese.
	}
\end{quote}

\noindent
This equated to approximately 1 in every 14 people in the world classed as obese in 2014, and therefore, 1 in 14 people around the world were also at severe health risks associated with obesity.
The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese, respectively \citep{WHO2014}.
It was also noted in the report that women were more likely to be obese than men, and that the prevalence of overweight and obesity increased as the income level of the countries increased \citep{WHO2014}.

In addition to this, the worldwide prevalence of childhood overweight and obesity has been increasing steadily since 2000 \citep{WHO2014}.
Furthermore, the prevalence of overweight in children under 5 years was estimated to reach from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
The current obesity status of the adult population is bad already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population in the future.

In \gls{nz}, obesity is a serious problem as approximately one third (31.6\%) of the adults in \gls{nz} were classified as obese in 2015/2016 \citep{Health2016}.
The Pacific and M\=aori populations were affected the most, where 67\% and 47\% of the adults, respectively, were obese \citep{Health2016}.
\gls{nz} not only have a high prevalence of obesity compared to many of the other countries of the world, but also have a unique health problem that affects the Pacific and M\=aori populations more than the rest of the \gls{nz} population.

@@

Obesity is caused by continuous intake of food without expending all of the energy gained from the food.
The key to the problem is the control of food intake and energy expenditure, and how they are regulated in the body.
The role of hypothalamus is central to the control of satiety and many of the mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
The hypothalamus receives both neuronal and hormonal inputs from the body and propagates the signal to the downstream neurons that ultimately affect the food intake of the individual \citep{Bell2005, Spiegelman2001}.

@@

Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
Briefly, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.

The exact role of insulin in the satiety control was not particularly clear even though there was evidence of increased food intake with neuron-specific deletion of insulin receptor \citep{Barsh2002,Bruning2000}.
The fact that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell: leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
The activation of K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.

Satiety control is crucial in order to manage  and maintain the body weight in a healthy state.
As described above, there are many components that regulate satiety and disruption of any part of this satiety controlling pathway can encourage the development of obesity.

@@

There have been many studies that have investigated the association of genes and/or genomic regions with obesity.
Of those, many of the \glspl{gwas} published in 2007 showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity  \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
Until these studies came out, there has been no common genes or genomic regions that associated with obesity that was reproducible and reliable \citep{Frayling2007}.
Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.

@@

It is clear that the individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
For this reason, obesity is thought to be caused mainly by the environmental factors that act on a population-level.
The first key factor that contributes to the global rise in the proportion of the population that is obese is the availability and abundance of certain food in the country.
With the help from the government trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
This led to greater abundance of food in general, greater number of choices of food and shifted the overall nutritional status of the country \citep{Malik2013}.
This shift in the food choice and availability (and the economic benefit associated with the trades) had a positive impact on the country, but the growth in the food market also resulted in the population-wide rise in obesity.

As an example, in the \gls{usa}, the costs of corn and soy are low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
As a result, more people are inclined to consume food that are cheap, have little nutritional value and high in energy, than the food that are expensive, nutritional and healthy alternatives.
Thus, the population is more likely to become obese by making poor food choice.

The second factor to consider is the behavioural changes led by the urbanisation of the population.
As more people take up urban lifestyle, they are exposed to the new environment where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also impose negative health behaviour that ultimately lead to positive energy balance and therefore obesity \citep{Malik2013}.
Lack of physical activity is one of the obvious components that is linked with the urbanisation of the population.
In many urban cities, the recommended level of physical exercise is often difficult to achieve and maintain due to the shift towards the sedentary behaviours that are encouraged by the developed transport system, automated household chores and indoor entertainment \citep{Malik2013}.

The last factor to be mentioned here is the income and socioeconomic status of the population.
As the average income increases, more people are able to buy food and are likely to take up the sedentary lifestyle.
From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary to this expectation, the biggest impact from obesity is seen in the low- and middle-income groups of the population \citep{Malik2013}.
Access to the developed health care system will allow weight maintenance of the population, but this is likely to be of most benefit for the high-income group in the population.
The reason for this, quite simply, is the cost associated with the visit to the health care system, as high-income group is more likely to make a consistent and periodic visit than the low- or middle-income groups of the population \citep{Malik2013}.

So, for the low- and middle-income groups, the rise in the average income in these groups allow the people to have access to food and activities that promotes sedentary behaviours, but the limited access to the health care system, education and recreational activities that promote physical activities ultimately lead to positive energy balance \citep{Malik2013}.
In contrast, the high-income group of the population have better access to the facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
All of these population-level environmental factors in synergy with the individual-level genetic factors have contributed to the global epidemic of obesity.

@@

\citet{Bray2013} estimated the global prevalence of cancer in adult population in 2008.
Since there was no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer, with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
When the population of these regions was taken into account and the proportion of cancer prevalence calculated, Western Europe had the highest prevalence proportion of 1.9\%, followed closely by Australia/New Zealand with 1.82\% and North America with 1.7\% \citep{Bray2013}.

Looking at the prevalence of cancer by specific types, breast cancer was the most prevalent type of cancer with 5.2 million cases in 2008 \citep{Bray2013}.
Colorectal and prostate cancers were second and third most prevalent cancer types, with approximately 3.2 million cases for  both of them \citep{Bray2013}.
\citet{Bray2013} noted that female-specific cancer types were the most prevalent types of cancer in 75\% of the countries in the world.
Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of the countries \citep{Bray2013}.
Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of the prevalence of cancer  around the world.

In \gls{nz}, the most common cancer types diagnosed for males are prostate, colorectal, melanoma, lung and non-Hodgkin lymphoma, which all together accounted for 65\% of newly registered cancers in males in 2013 \citep{Health2016b}.
For females, breast (which accounted for 28.3\% of all new female cancer cases), colorectal, melanoma, lung and uterine cancers were the most common cancer types diagnosed in 2013 \citep{Health2016b}.
The cancer registration rates were higher in females aged 25-54 years than males of the same age group, but the registration rates were significantly higher in males for people who were 60 years or older \citep{Health2016b}.
Like with obesity, there is a significant difference in the cancer registration rates between the M\=aori and the non-M\=aori populations.
The rates were 1.2 and 1.4 times greater in the M\=aori population compared to the non-M\=aori population for males and females, respectively \citep{Health2016b}.
Again, this difference between the M\=aori and non-M\=aori population is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes for the M\=aori population.

The total number of deaths by non-communicable diseases (including cancer) is predicted to increase, reaching 52 million deaths by 2030 \citep{WHO2014}.
With the prevalence of cancer at such high levels around the world, and with cancer mortality increasing in the future, cancer is a serious burden that the world has to face.

@@

There are two barriers in achieving infinite replicative capacity: senescence and crisis \citep{Hanahan2011}.
When cells are grown in a culture, cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
The cells that manage to bypass this senescence phase by mutating the tumour suppressor proteins such as the p53 protein, undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
Very rarely, some cells are able to survive the crisis phase and acquire the ability to replicate infinitely, and this transition has been termed ``immortalisation'' \citep{Hanahan2011, Wright1989}.
Without immortalisation, cancer cells are not able to generate a macroscopic tumours that are capable of killing the host organism \citep{Hanahan2000,Hanahan2011}.

@@

\noindent
Like normal cells, cancer cells also require access to oxygen and nutrient supplies from the blood.
However, new blood vessel formation, or angiogenesis, only occur in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.

It is thought that the angiogenic ability of the tumour is gained at an early stage of tumour development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
\Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have inhibitory effect) \citep{Yancopoulos2000}.
Opposing the effect of \Gls{vegfa} is the \gls{tsp1}, a protein that is known to inhibit the process of angiogenesis by affecting the bioavailability of \acrshort{vegf}, as well as promoting apoptosis \citep{Kazerounian2008}.
In general, many cancers stimulate the expression of \gls{vegfa} protein and reduces the expression of \gls{tsp1} \citep{Kazerounian2008}.

@@

\noindent
One of the prime reasons that cancers are so difficult to treat and have high mortality rate is due to their ability to metastasise to other parts of the body.
Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology as the primary tumour, and thus have to consider multiple treatment plans to cure the patient.

Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
Once mobile and invasive, metastatic cancer cell is able to circulate in the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
There are numerous signals, growth factors and pathways that initiate the metastatic phenotype in cancer cell, and it remains an active field of research \citep{Hanahan2011,Kalluri2009}.

@@

Until recently, the evolutionary progress of cancers have been thought to be gradual and stepwise, where each driver mutation is acquired over many years, or even decades \citep{Stephens2011}.
There are now evidence that show otherwise, whereby a single catastrophic event (termed \gls{chromothripsis}) occurs in the chromosome structure, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
Owing to its nature of being a one-off chromosomal disruption, \gls{chromothripsis} (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.

In addition to this, areas of the chromosome where major structural rearrangements have occurred can be prone to a regional hypermutation, termed \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
Although the full mechanism in which \gls{kataegis} occur is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately encourage the acquisition of many of the tumorigenic phenotypes.

@@

\Gls{immunoediting} is a process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells and therefore poses a selective advantage on these cells \citep{Hanahan2011,Teng2008}.
In fact, there have been reports where the cancer cells in the immunodeficient mice not being able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in the immunodeficient mice \citep{Hanahan2011}.
One possible reason for this was that the cancer cells from the immunodeficient mice were not immunoedited to the extent in which the cells from immunocompetent mice were, and thus unable to evade the immune system as efficiently as those cells in the immunocompetent mice \citep{Hanahan2011}.
Though more experiments must be carried out to confirm its definitive role and mechanisms in which immunoediting contributes to cancer biology, immune system evasion is no doubt a significant factor that contribute to tumour progression. \\

\noindent
By utilising many, if not all, of these major cancer hallmarks, cancer cells are able to initiate growth, proliferate, invade, metastasise and eventually kill the host if untreated.
There is an evident need for an efficient way to monitor and treat cancer patients, and these Hallmarks of Cancers will prove to be the best target for effective treatments of cancer in the future.

@@

Cancers are developed through many cellular pathways (Hallmarks of Cancers) that are not fully uncovered yet, with either too many or very few causes associated with those pathways.
Therefore it is difficult to assess how, rather than whether, a given environmental factor causes specific cancer.
With that said, several environmental factors have a strong evidence of causing cancer.

>>>>

Obesity has been a major global problem for more than a decade, associated with many noncommunicable diseases such as diabetes, cardiovascular diseases and certain types of cancers \citep{WHO2014}.
In fact, the risk of comorbidities increases with an increase in \gls{bmi}, where the risk becomes severe as the \gls{bmi} level approaches the obese category \citep{WHO2000}.
The number of overweight and obese people, in both adults and children, has risen in every country of the world and the trend will likely to continue in the future.
It is estimated to account for 3.4 million deaths per year and 93.6 million \glspl{daly} in 2010, obesity is a serious disease that continues to grow in our society \citep{Lim2012}.

@@

Obesity is defined as an abnormal or excessive fat accumulation that may impair the health of an individual \citep{Garrow1988}.
One common and widely used approach to categorise obesity is to measure the \gls{bmi} of the individual.
\gls{bmi} is a measurement based on the weight-to-height ratio of an individual.
It is often used by clinicians and epidemiologists to classify adults into underweight, normal weight, overweight, or obese categories.
The unit of \gls{bmi} is defined as the weight in kilograms per square of the height in metres (kg/m$^2$).

The \gls{who} (2014) categorise overweight and obese adults  with a \gls{bmi} of  $\geq$ 25 kg/m$^2$ and $\geq$ 30 kg/m$^2$ respectively.
For children under the age of 5, overweight and obesity is categorised as a  weight-for-height ratio greater than 2 standard deviation and 3 standard deviation above the \gls{who} Child Growth Standards median respectively.
For children aged between 5 to 19 years, overweight and obesity is defined as \gls{bmi}-for-age greater than 1 standard deviation and 2 standard deviation above the \gls{who} Growth Reference Standards median respectively.

However, the use of \gls{bmi} as an accurate measure of body fat composition and/or representation of the metabolic state of an individual remains controversial.
In fact, many studies have suggested that the use of other measurements, such as waist-to-height and waist-to-hip ratios, to classify the patients into different categories, as these measurements were better predictors of the diseases related to obesity than \gls{bmi} \citep{Dalton2003,Gelber2008,Lee2008}.
With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements.
As a result, most of the publicly available clinical data has height and weight data of the patients from which the \gls{bmi} can be calculated.
% With that said, \gls{bmi} is still widely used due to the ease of data collection compared to other measurements and, as a result, most of the publicly available clinical data have  height and weight data of the patients from which one can calculate the \gls{bmi} information.

@@

From the latest global status report on noncommunicable diseases by the \citet{WHO2014}:
\begin{quote}
	\textit{
		``In 2014, 39\% of adults aged 18 years and older (38\% of men and 40\% of women) were overweight.
		The worldwide prevalence of obesity nearly doubled between 1980 and 2014.
		In 2014, 11\% of men and 15\% of women worldwide were obese.
		Thus, more than half a billion adults worldwide are classed as obese.''
	}
\end{quote}

\noindent
Approximately 1 in every 14 people in the world were classed as obese in 2014; therefore, 1 in every 14 people around the world were also at severe health risks associated with obese people.
The highest prevalence of overweight and obesity was in the American regions, where 61\% and 27\% of the population were overweight and obese respectively \citep{WHO2014}.
It was also noted in the report that women were more likely to be obese than men, and that the prevalence of overweight and obesity increased as the income level of the countries increased \citep{WHO2014}.

In addition to this, the worldwide prevalence of childhood obesity has been increasing steadily since 2000 \citep{WHO2014}.
The prevalence of overweight children in those under 5 was estimated to rise from 6.3\% in 2013 to 11\% by 2025 if the current trend continued \citep{WHO2014}.
% The current obesity status of the adult population is high already, but the fact that more of the younger population who are overweight and obese poses a much greater risk for the health of the population in the future.
The current prevalence of obesity in the younger population poses an even greater risk to the health of the population in the future than the current prevalence of obesity in the adult population.

In \gls{nz}, obesity is a serious problem as approximately one third (31.6\%) of adults in \gls{nz} were classified as obese in 2015/2016 \citep{Health2016}.
The Pacific and M\=aori populations were affected the most, where 67\% and 47\% of the adults, respectively, were obese \citep{Health2016}.
\gls{nz} not only has a high prevalence of obesity compared to many other countries of the world, but also has a unique health problem that affects Pacific and M\=aori populations more than the rest of the \gls{nz} population.

@@

Obesity is caused by continuous intake of food without expending all of the energy gained from the food.
The key to the problem is the control of food intake and energy expenditure, and how they are regulated in the body.
The role of the hypothalamus is central to the control of satiety and many of the mutations that cause obesity affects the satiety signals to and/or from hypothalamus.
The hypothalamus receives both neuronal and hormonal inputs from the body and propagates the signal to the downstream neurons which ultimately affects the food intake of the individual \citep{Bell2005, Spiegelman2001}.

@@

Insulin, perhaps the most well-known hormone for its role in glucose homeostasis, also has an important role in satiety and food intake regulation.
In brief terms, when insulin binds to the insulin receptor, \gls{irs} protein family is recruited and phosphorylated by the  intrinsic tyrosine kinase activity of the insulin receptor \citep{Saltiel2002}.
The phosphorylated \gls{irs} protein then activates \gls{pi3k}, which in turn activates its downstream targets, causing a rapid signal cascade through the cell \citep{Saltiel2002}.

The exact role of insulin in the satiety control was not particularly clear even though there was evidence of increased food intake with neuron-specific deletion of insulin receptor \citep{Barsh2002,Bruning2000}.
The fact that the insulin level, like leptin, was proportional to the fat mass also suggested the role of insulin in satiety and body weight control \citep{Barsh2002, Bruning2000, Woods1979}.
This was perhaps due to the fact that leptin and insulin used two different signalling pathways in the cell: leptin used the \gls{jak}/\gls{stat} signalling pathway, whereas insulin used the \gls{pi3k} pathway for signalling \citep{Ghilardi1996}.
However, leptin, and subsequently insulin, was shown to activate ATP-dep\-endent K$^+$ channel in glucose-responsive neurons in the hypothalamus \citep{Spanswick1997, Spanswick2000}.
The activation of the K$^+$ channel provided evidence for the two pathways to act similarly on the cell, at least for an acute response, suggesting a convergence in the pathway response by the two hormones, and therefore their role in satiety control.

Satiety control is crucial in order to manage  and maintain the body weight in a healthy state.
As described above, there are many components that regulate satiety.
Disruption of any part of the satiety controlling pathway can encourage the development of obesity.

@@

There have been many studies that have investigated the association of genes and/or genomic regions with obesity.
Of those, many of the \glspl{gwas} published in 2007 showed strong evidence of the \gls{fto} gene with \gls{bmi} and obesity  \citep{Dina2007,Frayling2007,Gerken2007,Scuteri2007}.
Until these studies came out, there has been no common genes or genomic regions associated with obesity that was reproducible and reliable \citep{Frayling2007}.
Thus, these studies were the first evidence of common variants that contributed to the polygenic obesity phenotype in the population.

@@

It is clear that the individual-level factors cannot be accounted for the global epidemic of obesity that the world currently observes.
For this reason, obesity is thought to be caused mainly by the environmental factors that act on a population-level.
The first key factor that contributes to the global rise in the proportion of the population that is obese is the availability and abundance of certain foods in the country.
With the help from government trade policies, food and other goods have become easier to trade between countries, helping the economic growth of many countries around the world \citep{Kearney2010}.
This led to a greater abundance of food in general, a greater number of food choices and shifted the overall nutritional status of countries \citep{Malik2013}.
This shift in food choice and availability (and the economic benefit associated with the trades) had a positive impact on countries, but the growth in the food market also resulted in the population-wide rise in obesity.

As an example, in the \gls{usa}, the costs of corn and soy are low due to the fact that they are the raw ingredients for most processed food and beverages, such as high fructose corn syrups used to sweeten many soft drinks in the \gls{usa} \citep{Malik2013}.
Moreover, corn and soy are also the main food for the livestock, which results in lower prices for meat \citep{Malik2013}.
On the other hand, the prices of fruits and vegetables remain expensive due to the lack of support by the government to lower the cost associated with the production \citep{Malik2013}.
As a result, more people are inclined to consume foods that are cheap, have little nutritional value and high in energy, than foods that are expensive, nutritional and healthy alternatives.
Thus, the population is more likely to become obese by making poor food choice.

The second factor to consider is the behavioural changes led by urbanisation.
As more people take up urban lifestyle, they are exposed to new environments where there is a better range of food selection, as well as technological and mechanical advancement in transport and other daily chores that may not be available in rural areas.
Although there are many advantages involved with urbanisation, such as access to developed health care systems, education and advanced technologies, the lifestyle change also imposes negative health behaviour that ultimately can lead to positive energy balance and obesity \citep{Malik2013}.
Lack of physical activity is one of the obvious components linked with urbanisation.
In many urban cities, the recommended level of physical exercise is often difficult to achieve and maintain due to the shift towards sedentary behaviours that are encouraged by developed transport systems, automated household chores and indoor entertainment \citep{Malik2013}.

The last factor to be mentioned here is the income and socioeconomic status of the population.
As the average income increases, more people are able to buy food and are likely to take up a sedentary lifestyle.
From this, one would expect the high-income group to have the largest impact from obesity, but on the contrary, the biggest impact from obesity is seen in the low- and middle-income groups \citep{Malik2013}.
Access to a developed health care system will allow weight maintenance, but this is likely to be of most benefit to high-income groups.
The reason for this is the cost associated with a visit to the health care system, as high-income groups are more likely to make a consistent and periodic visit than a low- or middle-income groups \citep{Malik2013}.

So, for the low- and middle-income groups, the rise in the average income in these groups allows people to have access to food and activities that promotes sedentary behaviours, but the limited access to the health care system, education and recreational activities (that promote physical activity) ultimately leads to positive energy balance \citep{Malik2013}.
In contrast, the high-income group of the population have better access to facilities that promote weight maintenance, thus have less of an impact compared with the low- and middle-income groups of the population.
All of these population-level environmental factors in synergy with the individual-level genetic factors have contributed to the global epidemic of obesity.

@@

\citet{Bray2013} estimated the global prevalence of cancer in the adult population in 2008.
Since there was no clear definition for cancer prevalence, \citet{Bray2013} used \gls{ldp} as an estimate for cancer prevalence, and measured the incidence of cancer during the period between 2004 and 2008 (5 years).
From these estimates, Eastern Asia had the greatest 5-year prevalence of cancer, with approximately 7 million people with cancer in 2008, followed by North America and Western Europe with approximately 4.7 and 3 million people with cancer \citep{Bray2013}.
When the population of these regions was taken into account and the proportion of cancer prevalence calculated, Western Europe had the highest prevalence proportion of 1.9\%, followed closely by Australia/New Zealand with 1.82\% and North America with 1.7\% \citep{Bray2013}.

Looking at the prevalence of cancer by specific types, breast cancer was the most prevalent type of cancer with 5.2 million cases in 2008 \citep{Bray2013}.
Colorectal and prostate cancers were the second and third most prevalent cancer types, with approximately 3.2 million cases \citep{Bray2013}.
\citet{Bray2013} noted that female-specific cancer types were the most prevalent types of cancer in 75\% of countries in the world.
Furthermore, the prevalence of breast, cervix or prostate cancer is the highest in almost 95\% of countries \citep{Bray2013}.
Although these estimates provide a relatively crude measure of cancer prevalence, \citet{Bray2013} provided a useful overview of the prevalence of cancer  around the world.

In \gls{nz}, the most common cancer types diagnosed for males are prostate, colorectal, melanoma, lung and non-Hodgkin lymphoma, which all together accounted for 65\% of newly registered cancers in males in 2013 \citep{Health2016b}.
For females, breast (which accounted for 28.3\% of all new female cancer cases), colorectal, melanoma, lung and uterine cancers were the most common cancer types diagnosed in 2013 \citep{Health2016b}.
The cancer registration rates were higher in females aged 25-54 years than males of the same age group, but the registration rates were significantly higher in males for people who were 60 years or older \citep{Health2016b}.
Like with obesity, there is a significant difference in the cancer registration rates between M\=aori and non-M\=aori.
The rates were 1.2 and 1.4 times greater in the M\=aori population compared to the non-M\=aori population for males and females, respectively \citep{Health2016b}.
Again, this difference between the M\=aori and non-M\=aori is unique to \gls{nz}, and improved understanding of the underlying problem is essential for better treatment plans and outcomes for M\=aori.

The total number of deaths by non-communicable diseases (including cancer) is predicted to increase, reaching 52 million deaths by 2030 \citep{WHO2014}.
With the prevalence of cancer at such high levels around the world, and with cancer mortality increasing in the future, cancer is a serious challenge that the world has to face.

@@

There are two barriers in achieving infinite replicative capacity: senescence and crisis \citep{Hanahan2011}.
When cells are grown in a culture, cells are able to replicate and grow until a certain point, and the cells reach the stage of senescence where the cells stop growing but are still viable \citep{Hanahan2011}.
The cells that manage to bypass this senescence phase by mutating the tumour suppressor proteins such as the p53 protein, undergoes the crisis phase where the majority of the cells in the culture dies \citep{Hanahan2011}.
Very rarely, some cells are able to survive the crisis phase and acquire the ability to replicate infinitely, and this transition has been termed ``immortalisation'' \citep{Hanahan2011, Wright1989}.
Without immortalisation, cancer cells are not able to generate macroscopic tumours that are capable of killing the host organism \citep{Hanahan2000,Hanahan2011}.

@@

\noindent
Like normal cells, cancer cells also require access to oxygen and nutrient supplies from the blood.
However, new blood vessel formation, or angiogenesis, only occurs in a handful of occasions, and thus the tumour needs to acquire the ability to stimulate angiogenesis in the microenvironment in which they live in \citep{Hanahan2011}.

It is thought that the angiogenic ability of the tumour is gained at an early stage of tumour development by disrupting the ``angiogenic switch'' \citep{Hanahan2011}.
\Gls{vegfa} is the major growth factor that triggers angiogenesis by signalling through the \acrshort{vegf} receptors (though some receptors have an inhibitory effect) \citep{Yancopoulos2000}.
Opposing the effect of \Gls{vegfa} is the \gls{tsp1}, a protein that is known to inhibit the process of angiogenesis by affecting the bioavailability of \acrshort{vegf}, as well as promoting apoptosis \citep{Kazerounian2008}.
In general, many cancers stimulate the expression of \gls{vegfa} protein and reduces the expression of \gls{tsp1} \citep{Kazerounian2008}.

@@

\noindent
One of the prime reasons that cancers are so difficult to treat and have a high mortality rate is due to their ability to metastasise to other parts of the body.
Once metastasised to different parts of the body, the clinician not only has to consider the primary tumour site, but also the second, third, or even fourth tumour sites, which may or may not have similar tumour biology as the primary tumour, and thus has to consider multiple treatment plans to cure the patient.

Metastasis is a multi-step process that requires the selection and survival of cancer cell(s) that is able to metastasise (``seed''), and is also compatible with the specific target tissue where they prosper (``soil'') \citep{Talmadge2010}.
Once mobile and invasive, the metastatic cancer cell is able to circulate in the vascular system and eventually invade the target organ, where it must undergo \gls{met} and colonise the site \citep{Hanahan2011,Kalluri2009}.
There are numerous signals, growth factors and pathways that initiate the metastatic phenotype in the cancer cell, and it remains an active field of research \citep{Hanahan2011,Kalluri2009}.

@@

Until recently, the evolutionary progress of cancers have thought to have been gradual and stepwise, where each driver mutation is acquired over many years, or even decades \citep{Stephens2011}.
There is now evidence that shows otherwise, whereby a single catastrophic event (termed \gls{chromothripsis}) occurs in the chromosome structure, resulting in major structural rearrangements such as inversions, deletions, duplications and translocations \citep{Leibowitz2015,Stephens2011}.
Owing to its nature of being a one-off chromosomal disruption, \gls{chromothripsis} (usually) cause disruptions in tumour suppressor genes rather than duplications of oncogenes \citep{Leibowitz2015}.

In addition to this, areas of the chromosome where major structural rearrangements have occurred can be prone to a regional hypermutation, termed \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
Although the full mechanism in which \gls{kataegis} occurs is not yet defined, \gls{aid}/\acrshort{apobec} family proteins, a family of proteins involved in gene editing, may play a major role in \gls{kataegis} \citep{Leibowitz2015,Nik-Zainal2012}.
It is evident that both \gls{chromothripsis} and \gls{kataegis} contribute to the genomic aberrations and mutations that ultimately encourage the acquisition of many of the tumorigenic phenotypes.

@@

\Gls{immunoediting} is a process where the immune system successfully eliminates the highly immunogenic cancer cells during surveillance, but fails to eliminate the weakly immunogenic cells and therefore poses a selective advantage on these cells \citep{Hanahan2011,Teng2008}.
In fact, there have been reports where the cancer cells in the immunodeficient mice were not able to initiate tumour growth in the immunocompetent mice, but the cancer cells from the immunocompetent mice were able to do so in the immunodeficient mice \citep{Hanahan2011}.
One possible reason for this was that the cancer cells from the immunodeficient mice were not immunoedited to the extent in which the cells from immunocompetent mice were, and thus unable to evade the immune system as efficiently as those cells in the immunocompetent mice \citep{Hanahan2011}.
Though more experiments must be carried out to confirm its definitive role and mechanisms in which immunoediting contributes to cancer biology, immune system evasion is no doubt a significant factor that contributes to tumour progression. \\

\noindent
By utilising many, if not all, of these major cancer hallmarks, cancer cells are able to initiate growth, proliferate, invade, metastasise and eventually kill the host if untreated.
There is an evident need for an efficient way to monitor and treat cancer patients, and these Hallmarks of Cancers will prove to be the best target for effective treatment of cancer in the future.

@@

Cancers are developed through many cellular pathways (Hallmarks of Cancers) that are not fully uncovered yet, with either too many or very few causes associated with those pathways.
Therefore it is difficult to assess how, rather than whether, a given environmental factor causes a specific cancer.
With that said, several environmental factors have a strong evidence of causing cancer.
